\LaTeXe\ guide for the Journal of Functional Programming

### 2019-04-28 papers/jfp/5-rebuilders.tex
Minor tweaks to S5

The most complex argument of \hs{verifyVT} is a function
\hs{fetchHash}~\hs{::}~\hs{k}~\hs{->}~\hs{m}~\hs{(Hash}~\hs{v)} to obtain the
hash of the post-build value of any key. With an applicative task,
\hs{fetchHash} will be called on the statically known task dependencies.
However, with a monadic task, the dependencies are not known from the task
alone, they are only recorded from previous executions and stored in \hs{VT}. If
the build system has two traces for a given key~\hs{k}, they will both request
the same dependency first, since \hs{Task}~\hs{Monad} is deterministic. However,
depending on that first result, they may then request different successive
dependencies using \hs{fetchHash}. A curious result is that for suspending
schedulers~(\S\ref{sec-suspending}) in many cases the actual build steps are
performed as a consequence of checking if a key needs rebuilding!

@@

The \Shake build system and the associated paper -- see~\S2.3.3
in~Mitchell~\shortcite{mitchell2012shake} -- use a slightly different trace
structure, called \emph{step traces}, one that is similar to verifying traces,
but stores less data, and has different early cutoff semantics. Rather than
storing the \hs{Hash}~\hs{v} for each dependency, it instead stores \hs{built}
time and \hs{changed} time for the \hs{key}, and a list of dependency keys
(without the hashes). The resulting \hs{StepTrace} type resembles:

@@

This optimisation is useful, at least in the case of \Shake, to save space. A
typical cryptographic hash takes up 32 bytes, while a key (in \Shake only) is an
\hs{Int} taking only~4~bytes. Furthermore, \Shake permits values to be
arbitrarily large, and permits a custom equality (two values can be bit-for-bit
unequal but considered equal by \Shake), meaning that a \hs{Hash} is not a valid
encoding. For build systems that use \hs{Applicative} dependencies, the
\hs{depends} could be omitted entirely, making the size of a \hs{StepTrace}
$O(1)$ instead of $O(n)$, where $n$ is the number of dependencies.

>>>>

The most complex argument of \hs{verifyVT} is a function
\hs{fetchHash}~\hs{::}~\hs{k}~\hs{->}~\hs{m}~\hs{(Hash}~\hs{v)} to obtain the
hash of the post-build value of any key. With an applicative task,
\hs{fetchHash} will be called on the statically known task dependencies.
However, with a monadic task, the dependencies are not known from the task
alone, they are only recorded from previous executions stored in \hs{VT}. If
the build system has two traces for a given key~\hs{k}, they will both request
the same dependency first, since \hs{Task}~\hs{Monad} is deterministic. However,
based on that first result, they may then request different subsequent
dependencies using \hs{fetchHash}. A curious result is that for suspending
schedulers~(\S\ref{sec-suspending}) in many cases the actual build steps are
performed as a consequence of checking if a key needs rebuilding!

@@

The \Shake build system and the associated paper -- see~\S2.3.3
in~Mitchell~\shortcite{mitchell2012shake} -- use a slightly different trace
structure, called \emph{verifying step traces},
which stores less data than veryifying traces, and has slightly different early cutoff semantics. Rather than
storing the \hs{Hash}~\hs{v} for each dependency, it instead stores \hs{built}
time and \hs{changed} time for each \hs{key}, and a list of dependency keys
(without the hashes). The resulting \hs{StepTrace} type resembles:

@@

This optimisation is useful, at least in the case of \Shake, to save space. A
typical cryptographic hash takes up 32 bytes, while a key (in \Shake) is an
\hs{Int} taking only~4~bytes. Furthermore, \Shake permits values to be
arbitrarily large, and permits a custom equality (two values can be bit-for-bit
unequal but considered equal by \Shake), meaning that a \hs{Hash} is not a valid
encoding. For build systems that use \hs{Applicative} dependencies, the
\hs{depends} could be omitted entirely, making the size of a \hs{StepTrace}
$O(1)$ instead of $O(n)$, where $n$ is the number of dependencies.
